<<<<<<< HEAD
---
title: "Lab 8, Team All Stars"
author: "Johnny Tamanaha, Emily Archambault, Eugene Miller, Kaiyuan Fan, Jingzhi Zhang"
output: html_document
---
## Prep for Lab 9
Domain Expert: Sven Johannson (Economist at Stockholm School of Economics)

Team Question: What factors appear to have the greatest influence on the Happiness Score of a country?
This question is important because understanding the reasons behind what makes the people in some countries happier than others is the key to improving overall happiness for everyone. Finding what the common links are can guide how policies are developed, can shape our understanding of cultures around the world, and can give better insight into the biggest problems facing people today.

Dataset Description: The United Nations World Happiness Report is a survey of the state of global happiness. The 2017 report ranks 155 countries by their happiness levels using the happiness scores and rankings use data from the Gallup World Poll. This report estimates the extend to which six variables contribute to happiness score: GDP per Capita, Family, Life Expectancy, Freedom, Generosity, Trust, Government Corruption. Adding these columns up gives happiness score.

## Individual Sub Questions

# Emily Archambault

How does the extent to which the perception of corruption/trust in government's contribution to the overall happiness score vary by region of the world?
How does the extent to which family's contribution to the overall happiness score vary by region of the world?

These questions are interesting because the answers could give a sense as to what factors are most important in people's lives, and how those factors could depend on where they live. It could give insight into how different countries are run, and provide lessons, either positive or negative, for the rest of the world.

# Eugene Miller

How does happiness score vary by region of the world? 
Do the happiest countries have a low or high variance of individual scores for family, corruption, gdp per capita etc?

These questions are important because they give us insight into why happy countries are happy in historical context, and to whether a high amount of one factor leads to happiness or whether a low variability between scores leads to a high amount of happiness. Is it better for a country to be strong in one specific asepct or be more well rounded when it comes to quality of life?

## Graphs for EA Lake Data

# Tanganyika Data
```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(readxl)
Tang <- read_excel("East Africa Lake Data.xlsx",sheet=6)

Tang <- select(Tang, `Year AD...1`,`BSi (normalized)`,`Year AD...4`,`Charcoal (normalized)`,`Year`,TEX86)
Tang <- rename(Tang,year1=`Year AD...1`,BSi=`BSi (normalized)`, year2=`Year AD...4`,Charcoal=`Charcoal (normalized)`,year3=`Year`)
Tang_1 <- Tang %>%
  select(year1,BSi)
Tang_2 <- Tang %>%
  select(year2, Charcoal)
Tang_3 <- Tang %>%
  select(year3, TEX86)

Tang_1$year1 <- round(Tang_1$year1)
Tang_2$year2 <- round(Tang_2$year2)
Tang_3$year3 <- round(Tang_3$year3)

Tang_1_2 <- full_join(Tang_1,Tang_2,by=c("year1"="year2"))

Tang <- full_join(Tang_1_2,Tang_3,by=c("year1"="year3"))
Tang.skinny <- gather(Tang,key="measurement",value="value",BSi, Charcoal,TEX86)
Tang.skinny

ggplot(data=Tang.skinny)+
  geom_line(aes(x=year1,y=value,color = measurement))
```


# Biogenic Silica in Malawi Visualization
```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(readxl)

data <- read_excel("East Africa lake data.xlsx", sheet = "Malawi", skip = 1)

data <- data %>%
  subset(select = -c(4,5,6,7))
colnames(data)[1] <- "Years before 2002"
colnames(data)[2] <- "Calandar Year"
colnames(data)[3] <- "BSi (Location 1)"
colnames(data)[4] <- "BSi (Location 2)"

ggplot(data) +
  geom_point(mapping = aes(x = `Calandar Year`, y = `BSi (Location 1)`), color = "indianred4") +
  geom_point(mapping = aes(x = `Calandar Year`, y = `BSi (Location 2)`), color = "palegreen4") +
  geom_smooth(mapping = aes(x = `Calandar Year`, y = `BSi (Location 1)`), color = "indianred3", se = FALSE) +
  geom_smooth(mapping = aes(x = `Calandar Year`, y = `BSi (Location 2)` ),color = "palegreen3", se = FALSE) +
  labs(y = "BSi", x = "Year", title = "A Comparison Between Measurments in Biogenic Silica at Two Locations in Malawi")
```

# VOI Rainfall Visualization
```{r, echo=FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(readxl)
rain <- read_excel("East Africa lake data.xlsx", sheet = 8)
rain <- select(rain, `Year`, `Jan`, `Feb`, `Mar`, `Apr`, `May`, `Jun`, `Jul`, `Aug`, `Sep`, `Oct`, `Nov`, `Dec`)
rain <- gather(rain, `Jan`, `Feb`, `Mar`, `Apr`, `May`, `Jun`, `Jul`, `Aug`, `Sep`, `Oct`, `Nov`, `Dec`, key = "Month", value = "Rainfall")

ggplot(data = rain) + geom_smooth(mapping = aes(x= Year, y = Rainfall, color = Month), se = FALSE)
```

# CRU and MAM Rainfall Data
```{r, echo = FALSE, message = FALSE, warning = FALSE}
data <- read_excel("East Africa lake data.xlsx", sheet="CRU and MAM rainfall data")
data <- data %>%
  select(-c("...6", "...7", "...8"))
data_MAM <- data %>%
  gather("GPCC MAM Normalized", "CRU MAM Normalized", key="rainy Season", value="Global_Normalized_Rainfall") %>%
  separate("rainy Season", into=c("Dataset", "rainy Season"), sep=" ") %>%
  mutate("Rainy_Season" = "Short Rain") %>%
  select("Year", "Dataset", "Rainy_Season", "Global_Normalized_Rainfall")
data_OND <- data %>%
  gather("GPCC OND Normalized", "CRU OND Normalized", key="rainy Season", value="Global_Normalized_Rainfall") %>%
  separate("rainy Season", into=c("Dataset", "rainy Season"), sep=" ") %>%
  mutate("Rainy_Season" = "Long Rain") %>%
  select("Year", "Dataset", "Rainy_Season", "Global_Normalized_Rainfall")
data2 <- rbind(data_MAM, data_OND)
ggplot(data2) +
  geom_point(mapping=aes(x=Year, y=Global_Normalized_Rainfall, color=Rainy_Season, shape=Dataset), size=3) +
  geom_point(mapping=aes(x=Year, y=Global_Normalized_Rainfall, shape=Dataset), color="grey90", size=1.25) +
  labs(title="Normalized Rainfall from 1901 to 2010")+
  ylab("Global Normalized Rainfall") +
  scale_color_discrete(name="Rainy Season")
=======
---
title: "Lab 8, Team All Stars"
author: "Johnny Tamanaha, Emily Archambault, Eugene Miller, Kaiyuan Fan, Jingzhi Zhang"
output: html_document
---
## Prep for Lab 9
Domain Expert: Sven Johannson (Economist at Stockholm School of Economics)

Team Question: What factors appear to have the greatest influence on the Happiness Score of a country?
This question is important because understanding the reasons behind what makes the people in some countries happier than others is the key to improving overall happiness for everyone. Finding what the common links are can guide how policies are developed, can shape our understanding of cultures around the world, and can give better insight into the biggest problems facing people today.

Dataset Description: The United Nations World Happiness Report is a survey of the state of global happiness. The 2017 report ranks 155 countries by their happiness levels using the happiness scores and rankings use data from the Gallup World Poll. This report estimates the extend to which six variables contribute to happiness score: GDP per Capita, Family, Life Expectancy, Freedom, Generosity, Trust, Government Corruption. Adding these columns up gives happiness score.

## Individual Sub Questions

# Emily Archambault

How does the extent to which the perception of corruption/trust in government's contribution to the overall happiness score vary by region of the world?
How does the extent to which family's contribution to the overall happiness score vary by region of the world?

These questions are interesting because the answers could give a sense as to what factors are most important in people's lives, and how those factors could depend on where they live. It could give insight into how different countries are run, and provide lessons, either positive or negative, for the rest of the world.

# Eugene Miller

How does happiness score vary by region of the world? 
Do the happiest countries have a low or high variance of individual scores for family, corruption, gdp per capita etc?

These questions are important because they give us insight into why happy countries are happy in historical context, and to whether a high amount of one factor leads to happiness or whether a low variability between scores leads to a high amount of happiness. Is it better for a country to be strong in one specific asepct or be more well rounded when it comes to quality of life?

# Johnny Tamanaha

How does freedom affect happiness in each continent?
How does freedom affect happiness in each region?

These questions are interesting because freedom is such a defining characteristic or regions and countries. This could show evidence for an optimal level of freedom to maximize happiness for citizens. 

## Graphs for EA Lake Data

# Biogenic Silica in Malawi Visualization
```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(readxl)

data <- read_excel("East Africa lake data.xlsx", sheet = "Malawi", skip = 1)

data <- data %>%
  subset(select = -c(4,5,6,7))
colnames(data)[1] <- "Years before 2002"
colnames(data)[2] <- "Calandar Year"
colnames(data)[3] <- "BSi (Location 1)"
colnames(data)[4] <- "BSi (Location 2)"

ggplot(data) +
  geom_point(mapping = aes(x = `Calandar Year`, y = `BSi (Location 1)`), color = "indianred4") +
  geom_point(mapping = aes(x = `Calandar Year`, y = `BSi (Location 2)`), color = "palegreen4") +
  geom_smooth(mapping = aes(x = `Calandar Year`, y = `BSi (Location 1)`), color = "indianred3", se = FALSE) +
  geom_smooth(mapping = aes(x = `Calandar Year`, y = `BSi (Location 2)` ),color = "palegreen3", se = FALSE) +
  labs(y = "BSi", x = "Year", title = "A Comparison Between Measurments in Biogenic Silica at Two Locations in Malawi")
```

# VOI Rainfall Visualization
```{r, echo=FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(readxl)
rain <- read_excel("East Africa lake data.xlsx", sheet = 8)
rain <- select(rain, `Year`, `Jan`, `Feb`, `Mar`, `Apr`, `May`, `Jun`, `Jul`, `Aug`, `Sep`, `Oct`, `Nov`, `Dec`)
rain <- gather(rain, `Jan`, `Feb`, `Mar`, `Apr`, `May`, `Jun`, `Jul`, `Aug`, `Sep`, `Oct`, `Nov`, `Dec`, key = "Month", value = "Rainfall")

ggplot(data = rain) + geom_smooth(mapping = aes(x= Year, y = Rainfall, color = Month), se = FALSE)
```

# CRU and MAM Rainfall Data
```{r, echo = FALSE, message = FALSE, warning = FALSE}
data <- read_excel("East Africa lake data.xlsx", sheet="CRU and MAM rainfall data")
data <- data %>%
  select(-c("...6", "...7", "...8"))
data_MAM <- data %>%
  gather("GPCC MAM Normalized", "CRU MAM Normalized", key="rainy Season", value="Global_Normalized_Rainfall") %>%
  separate("rainy Season", into=c("Dataset", "rainy Season"), sep=" ") %>%
  mutate("Rainy_Season" = "Short Rain") %>%
  select("Year", "Dataset", "Rainy_Season", "Global_Normalized_Rainfall")
data_OND <- data %>%
  gather("GPCC OND Normalized", "CRU OND Normalized", key="rainy Season", value="Global_Normalized_Rainfall") %>%
  separate("rainy Season", into=c("Dataset", "rainy Season"), sep=" ") %>%
  mutate("Rainy_Season" = "Long Rain") %>%
  select("Year", "Dataset", "Rainy_Season", "Global_Normalized_Rainfall")
data2 <- rbind(data_MAM, data_OND)
ggplot(data2) +
  geom_point(mapping=aes(x=Year, y=Global_Normalized_Rainfall, color=Rainy_Season, shape=Dataset), size=3) +
  geom_point(mapping=aes(x=Year, y=Global_Normalized_Rainfall, shape=Dataset), color="grey90", size=1.25) +
  labs(title="Normalized Rainfall from 1901 to 2010")+
  ylab("Global Normalized Rainfall") +
  scale_color_discrete(name="Rainy Season")
>>>>>>> master
```